---
title: "History Through The Presidents' Words"
author: "Chisheng Li"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
library(stringr)
library(plyr)
library(dplyr)
library(magrittr)
library(mgcv)
library(tm)
library(proxy)
library(ggdendro)
library(ggplot2)
set.seed(100)
```

## Hierarchical Clustering

The Presidents are clustered based on the similarity of their State of the Union texts using the Jensen–Shannon divergence method. There are approximately 5 groups, and it is apparent that the presidents of similar eras are grouped together.

```{r, echo=FALSE, warning=FALSE}
## Load data
sotu <- readLines(file("sotu1790-2015.txt"))

breaks <- grep("\\*\\*\\*", sotu)

# Combine every SOTU speech into a single list, then split the speeches into separate lines by ".", "?" and "!"
combine = list()
for (i in 1:228){
  combine[[i]] = paste(sotu[(breaks[i]+6):breaks[i+1]-1], sep=" ", collapse = " ")
}
combine[[229]] <- paste(sotu[(breaks[229]+6):breaks[229+1]-1], sep=" ", collapse = " ")
sotuList <- sapply(combine, function(x) strsplit(x, "[\\?|\\.|\\!]"))
```

```{r, echo=FALSE, results="hide"}
speechToWords = function(sentences){
  # sentences is character vector of sentences for each speech
  s1=gsub("'","",sentences)
  # Remove all digits from the content
  s2=gsub("[[:digit:]]","",s1)
  # Remove [Applause]
  s3=gsub("(Applause.)","",s2)
  # Remove [Laughter]
  s4=gsub("(Laughter.)","",s3)

  # Turn characters to lower case.
  s5=tolower(s4)
  
  # Use strsplit to split the text up by blanks and punctuation
  s6=strsplit(s5,"([[:blank:]]|[[:punct:]])")
  # Unlist the return value 
  s7=unlist(s6)
  
  # Drop any empty word " "
  s8=s7[s7!=""]
  
  # return a character vector of all words in the speech
  return(s8)
}
```

```{r, echo=FALSE, results="hide"}
# Extract every word used in each SOTU speech into a list Words
sotuWords <- lapply(sotuList, speechToWords)

load("speechesDF.Rda")
```

```{r, echo=FALSE, results="hide"}
# Computing the distances between every SOTU speech
# Sort the list of words into unique terms
# Unlist the return value and use unique() to get the bag of words
# Alphabetize the bag of words and call it uniqueWords
# 23571 unique words total
uniqueWords <- sort(unique(unlist(sotuWords)))

# Create a word vector for every SOTU speech
vector <- rep(0, length(uniqueWords))
names(vector) <- uniqueWords

wordVector = lapply(sotuWords, function(x){
  vector[names(table(x))]=table(x)
  return(vector)
})

# From the word vector, create a matrix with columns corresponding to speeches and rows to words 
wordMatrix <- matrix(unlist(wordVector), ncol  = length(wordVector), byrow = FALSE)
```

```{r, echo=FALSE, results="hide"}
# Load compSJDist.R to compute the linguinstic distance between every president using the Jensen-Shannon metric
source("compSJDist.R")

# Create a matrix presMatrix by adding the word vectors for the speecbes made by the same president
pres <- unique(speechesDF$president)

presMatrix <- matrix(0, nrow=length(uniqueWords),ncol=length(pres))
for(i in 1:length(pres)){
  if(sum(speechesDF$president==pres[i])!=1){
    presMatrix[,i]=rowSums(wordMatrix[,speechesDF$president==pres[i]])
  }else{
    presMatrix[,i]=wordMatrix[,speechesDF$president==pres[i]]
  }
}
```

```{r, echo=FALSE, results="hide", message=FALSE}
# Compute the document frequency df by calculating the number of presidents that used every word
df <- list()
for(i in 1:length(uniqueWords)){
  df[i] = sum(presMatrix[i,]!=0)
}
df <- unlist(df)

# Calculate the distance between every president based on the similarity of words that they used in their SOTU speech
# The lower distance values denote higher similarity
presDist <- compSJDist(tf = presMatrix, df=df, terms=uniqueWords)
rownames(presDist) <- unique(pres)
colnames(presDist) <- unique(pres)
```

```{r, echo=FALSE, fig.height=8, fig.width=14}
# Plot a hierarchical clustering of the presidents' similarity
hc <- hclust(as.dist(presDist))
ggdendrogram(hc, rotate = FALSE, size = 2) + labs(title="SOTU Presidents Cluster Dendrogram")
```

## K-Means Clustering

The Calinski-Harabasz function suggests that 6 is the optimal number of K-Means clusters to group the presidents.

```{r, echo=FALSE, message=FALSE}
# Gather list of all document files to be processed.
source("parseText.R")

files <- list.files(path = './data/')

sotu <- rbind_all(alply(.data = files, .margins = 1, .fun = parseText))
```

```{r, echo=FALSE, message=FALSE}
# Clean the data
sotu_corpus <- Corpus(VectorSource(sotu$content)) %>%
    tm_map(x = ., FUN = PlainTextDocument) %>%
    tm_map(x = ., FUN = removePunctuation) %>%
    tm_map(x = ., FUN = removeNumbers) %>%
    tm_map(x = ., FUN = removeWords, stopwords(kind = 'en')) %>%
    tm_map(x = ., FUN = stripWhitespace)
```

```{r, echo=FALSE, message=FALSE}
# Create a Document Term Matrix, with the documents as the rows, the individual words along the columns, and a frequency count as the content.
doc_term <- DocumentTermMatrix(sotu_corpus)
doc_term$dimnames$Docs <- sotu$file_name

# Create a TF-IDF matrix that is a matrix of the same dimentions as the Document Term Frequency matrix, except each frequency has been normalized to the frequency of the term in the entire document.
tf_idf <- weightTfIdf(m = doc_term, normalize = TRUE)
tf_idf_mat <- as.matrix(tf_idf)
```

```{r, echo=FALSE, message=FALSE}
# Since K-Means relies on Euclidean Distance rather than Cosine Dissimilarity, it is first necesary to normalize the TF-IDF matrix.
tf_idf_norm <- tf_idf_mat / apply(tf_idf_mat, MARGIN = 1, FUN = function(x) sum(x^2)^0.5)

pca_comp <- prcomp(tf_idf_norm)
pca_rep <- data_frame(sotu_name = sotu$file_name, pc1 = pca_comp$x[,1], pc2 = pca_comp$x[,2])
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load chIndex.R to choose the optimal number of K-Means clusters using the Calinski-Harabasz function 
source("chIndex.R")
x1 <- pca_rep
x1 <- x1[-c(1)]

a1 = chIndex(x1, kmax=10)
k1.hat = a1$k[which.max(a1$ch)]
k1 = kmeans(x1, k1.hat, iter.max=100, nstart=10, algorithm="Lloyd")
plot(a1$k,a1$ch, xlab='K', ylab='CH(K)', type='b', main='K-Means Clustering : CH Index vs K' )
```

```{r, echo=FALSE, message=FALSE}
# CH index suggest choosing 6 clusters
km_clust <- kmeans(x = tf_idf_norm, centers = 6, iter.max = 100)
pca_rep <- data_frame(sotu_name = sotu$file_name, pc1 = pca_comp$x[,1],
                      pc2 = pca_comp$x[,2], clust_id = as.factor(km_clust$cluster))
```

K-Means clustering also shows that the presidents are grouped together by similar eras based on their PCA features.

```{r, echo=FALSE, fig.height=8, fig.width=12}
ggplot(data = pca_rep, mapping = aes(x = pc1, y = pc2, color = clust_id)) +
    scale_color_brewer(palette = 'Set1') +
    geom_text(mapping = aes(label = sotu_name), size = 2.5, fontface = 'bold') +
    labs(title = 'K-Means Cluster: 6 clusters on PCA Features', x = 'PCA Factor 1', 
         y = 'PCA Factor 2') +
    theme_grey() +
    theme(legend.position = 'right', legend.title = element_blank())
```

## Word Associations

The following word associations shows how Democratic and Republican presidents differed in their choice of terms in their State of the Union speeches. 

Democratic presidents when they discussed "freedom":
```{r, echo=FALSE}
load("gopTDM.Rda")
load("demTDM.Rda")

findAssocs(demTDM, "freedom", 0.06)
```
Republican presidents when they discussed "freedom":
```{r, echo=FALSE}
findAssocs(gopTDM, "freedom", 0.05)
```
Democratic presidents when they discussed "budget":
```{r, echo=FALSE}
findAssocs(demTDM, "budget", 0.06)
```
Republican presidents when they discussed "budget":
```{r, echo=FALSE}
findAssocs(gopTDM, "budget", 0.06)
```
Democratic presidents when they discussed "energy":
```{r, echo=FALSE}
findAssocs(demTDM, "energy", 0.11)
```
Republican presidents when they discussed "energy":
```{r, echo=FALSE}
findAssocs(gopTDM, "energy", 0.08)
```
Democratic presidents when they discussed "security":
```{r, echo=FALSE}
findAssocs(demTDM, "security", 0.05)
```
Republican presidents when they discussed "security":
```{r, echo=FALSE}
findAssocs(gopTDM, "security", 0.05)
```
Democratic presidents when they discussed "economy":
```{r, echo=FALSE}
findAssocs(demTDM, "economy", 0.05)
```
Republican presidents when they discussed "economy":
```{r, echo=FALSE}
findAssocs(gopTDM, "economy", 0.05)
```

## Text Hierachical Clustering

Term hierachical clustering shows how frequent terms appeared together in each president's speeches and highlights the presidents' policies that were presented to Congress and to the American people.

```{r, echo=FALSE}
load("gw1TDM.Rda")
load("ja2TDM.Rda")
load("tj3TDM.Rda")
load("jm4TDM.Rda")
load("jm5TDM.Rda")
load("jqa6TDM.Rda")
load("aj7TDM.Rda")
load("mvb8TDM.Rda")
load("jt10TDM.Rda")
load("jp11TDM.Rda")
load("zt12TDM.Rda")
load("mf13TDM.Rda")
load("fp14TDM.Rda")
load("jb15TDM.Rda")
load("al16TDM.Rda")
load("aj17TDM.Rda")
load("usg18TDM.Rda")
load("rbh19TDM.Rda")
load("ca21TDM.Rda")
load("gc22TDM.Rda")
load("bh23TDM.Rda")
load("wm25TDM.Rda")
load("tr26TDM.Rda")
load("wt27TDM.Rda")
load("ww28TDM.Rda")
load("wh29TDM.Rda")
load("cc30TDM.Rda")
load("hh31TDM.Rda")
load("fdr32TDM.Rda")
load("ht33TDM.Rda")
load("de34TDM.Rda")
load("jfk35TDM.Rda")
load("lbj36TDM.Rda")
load("rn37TDM.Rda")
load("gf38TDM.Rda")
load("jc39TDM.Rda")
load("rr40TDM.Rda")
load("ghb41TDM.Rda")
load("wc42TDM.Rda")
load("gwb43TDM.Rda")
load("bo44TDM.Rda")
```

```{r, echo=FALSE}
gw1TDM2 <- removeSparseTerms(gw1TDM, sparse = 0.955)
gw1M <- as.matrix(gw1TDM2)
gw1Matrix <- dist(scale(gw1M))
gw1Fit <- hclust(gw1Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
ja2TDM2 <- removeSparseTerms(ja2TDM, sparse = 0.955)
ja2M <- as.matrix(ja2TDM2)
ja2Matrix <- dist(scale(ja2M))
ja2Fit <- hclust(ja2Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
tj3TDM2 <- removeSparseTerms(tj3TDM, sparse = 0.96)
tj3M <- as.matrix(tj3TDM2)
tj3Matrix <- dist(scale(tj3M))
tj3Fit <- hclust(tj3Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jm4TDM2 <- removeSparseTerms(jm4TDM, sparse = 0.955)
jm4M <- as.matrix(jm4TDM2)
jm4Matrix <- dist(scale(jm4M))
jm4Fit <- hclust(jm4Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jm5TDM2 <- removeSparseTerms(jm5TDM, sparse = 0.96)
jm5M <- as.matrix(jm5TDM2)
jm5Matrix <- dist(scale(jm5M))
jm5Fit <- hclust(jm5Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jqa6TDM2 <- removeSparseTerms(jqa6TDM, sparse = 0.955)
jqa6M <- as.matrix(jqa6TDM2)
jqa6Matrix <- dist(scale(jqa6M))
jqa6Fit <- hclust(jqa6Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
aj7TDM2 <- removeSparseTerms(aj7TDM, sparse = 0.96)
aj7M <- as.matrix(aj7TDM2)
aj7Matrix <- dist(scale(aj7M))
aj7Fit <- hclust(aj7Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
mvb8TDM2 <- removeSparseTerms(mvb8TDM, sparse = 0.96)
mvb8M <- as.matrix(mvb8TDM2)
mvb8Matrix <- dist(scale(mvb8M))
mvb8Fit <- hclust(mvb8Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jt10TDM2 <- removeSparseTerms(jt10TDM, sparse = 0.96)
jt10M <- as.matrix(jt10TDM2)
jt10Matrix <- dist(scale(jt10M))
jt10Fit <- hclust(jt10Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jp11TDM2 <- removeSparseTerms(jp11TDM, sparse = 0.96)
jp11M <- as.matrix(jp11TDM2)
jp11Matrix <- dist(scale(jp11M))
jp11Fit <- hclust(jp11Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
zt12TDM2 <- removeSparseTerms(zt12TDM, sparse = 0.96)
zt12M <- as.matrix(zt12TDM2)
zt12Matrix <- dist(scale(zt12M))
zt12Fit <- hclust(zt12Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
mf13TDM2 <- removeSparseTerms(mf13TDM, sparse = 0.965)
mf13M <- as.matrix(mf13TDM2)
mf13Matrix <- dist(scale(mf13M))
mf13Fit <- hclust(mf13Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
fp14TDM2 <- removeSparseTerms(fp14TDM, sparse = 0.955)
fp14M <- as.matrix(fp14TDM2)
fp14Matrix <- dist(scale(fp14M))
fp14Fit <- hclust(fp14Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jb15TDM2 <- removeSparseTerms(jb15TDM, sparse = 0.965)
jb15M <- as.matrix(jb15TDM2)
jb15Matrix <- dist(scale(jb15M))
jb15Fit <- hclust(jb15Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
al16TDM2 <- removeSparseTerms(al16TDM, sparse = 0.97)
al16M <- as.matrix(al16TDM2)
al16Matrix <- dist(scale(al16M))
al16Fit <- hclust(al16Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
aj17TDM2 <- removeSparseTerms(aj17TDM, sparse = 0.97)
aj17M <- as.matrix(aj17TDM2)
aj17Matrix <- dist(scale(aj17M))
aj17Fit <- hclust(aj17Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
usg18TDM2 <- removeSparseTerms(usg18TDM, sparse = 0.97)
usg18M <- as.matrix(usg18TDM2)
usg18Matrix <- dist(scale(usg18M))
usg18Fit <- hclust(usg18Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
rbh19TDM2 <- removeSparseTerms(rbh19TDM, sparse = 0.965)
rbh19M <- as.matrix(rbh19TDM2)
rbh19Matrix <- dist(scale(rbh19M))
rbh19Fit <- hclust(rbh19Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
ca21TDM2 <- removeSparseTerms(ca21TDM, sparse = 0.97)
ca21M <- as.matrix(ca21TDM2)
ca21Matrix <- dist(scale(ca21M))
ca21Fit <- hclust(ca21Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
gc22TDM2 <- removeSparseTerms(gc22TDM, sparse = 0.97)
gc22M <- as.matrix(gc22TDM2)
gc22Matrix <- dist(scale(gc22M))
gc22Fit <- hclust(gc22Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
bh23TDM2 <- removeSparseTerms(bh23TDM, sparse = 0.97)
bh23M <- as.matrix(bh23TDM2)
bh23Matrix <- dist(scale(bh23M))
bh23Fit <- hclust(bh23Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
wm25TDM2 <- removeSparseTerms(wm25TDM, sparse = 0.97)
wm25M <- as.matrix(wm25TDM2)
wm25Matrix <- dist(scale(wm25M))
wm25Fit <- hclust(wm25Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
tr26TDM2 <- removeSparseTerms(tr26TDM, sparse = 0.975)
tr26M <- as.matrix(tr26TDM2)
tr26Matrix <- dist(scale(tr26M))
tr26Fit <- hclust(tr26Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
wt27TDM2 <- removeSparseTerms(wt27TDM, sparse = 0.97)
wt27M <- as.matrix(wt27TDM2)
wt27Matrix <- dist(scale(wt27M))
wt27Fit <- hclust(wt27Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
ww28TDM2 <- removeSparseTerms(ww28TDM, sparse = 0.97)
ww28M <- as.matrix(ww28TDM2)
ww28Matrix <- dist(scale(ww28M))
ww28Fit <- hclust(ww28Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
wh29TDM2 <- removeSparseTerms(wh29TDM, sparse = 0.975)
wh29M <- as.matrix(wh29TDM2)
wh29Matrix <- dist(scale(wh29M))
wh29Fit <- hclust(wh29Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
cc30TDM2 <- removeSparseTerms(cc30TDM, sparse = 0.975)
cc30M <- as.matrix(cc30TDM2)
cc30Matrix <- dist(scale(cc30M))
cc30Fit <- hclust(cc30Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
hh31TDM2 <- removeSparseTerms(hh31TDM, sparse = 0.975)
hh31M <- as.matrix(hh31TDM2)
hh31Matrix <- dist(scale(hh31M))
hh31Fit <- hclust(hh31Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
fdr32TDM2 <- removeSparseTerms(fdr32TDM, sparse = 0.975)
fdr32M <- as.matrix(fdr32TDM2)
fdr32Matrix <- dist(scale(fdr32M))
fdr32Fit <- hclust(fdr32Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
ht33TDM2 <- removeSparseTerms(ht33TDM, sparse = 0.975)
ht33M <- as.matrix(ht33TDM2)
ht33Matrix <- dist(scale(ht33M))
ht33Fit <- hclust(ht33Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
de34TDM2 <- removeSparseTerms(de34TDM, sparse = 0.97)
de34M <- as.matrix(de34TDM2)
de34Matrix <- dist(scale(de34M))
de34Fit <- hclust(de34Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jfk35TDM2 <- removeSparseTerms(jfk35TDM, sparse = 0.975)
jfk35M <- as.matrix(jfk35TDM2)
jfk35Matrix <- dist(scale(jfk35M))
jfk35Fit <- hclust(jfk35Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
lbj36TDM2 <- removeSparseTerms(lbj36TDM, sparse = 0.975)
lbj36M <- as.matrix(lbj36TDM2)
lbj36Matrix <- dist(scale(lbj36M))
lbj36Fit <- hclust(lbj36Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
rn37TDM2 <- removeSparseTerms(rn37TDM, sparse = 0.975)
rn37M <- as.matrix(rn37TDM2)
rn37Matrix <- dist(scale(rn37M))
rn37Fit <- hclust(rn37Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
gf38TDM2 <- removeSparseTerms(gf38TDM, sparse = 0.975)
gf38M <- as.matrix(gf38TDM2)
gf38Matrix <- dist(scale(gf38M))
gf38Fit <- hclust(gf38Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
jc39TDM2 <- removeSparseTerms(jc39TDM, sparse = 0.97)
jc39M <- as.matrix(jc39TDM2)
jc39Matrix <- dist(scale(jc39M))
jc39Fit <- hclust(jc39Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
rr40TDM2 <- removeSparseTerms(rr40TDM, sparse = 0.975)
rr40M <- as.matrix(rr40TDM2)
rr40Matrix <- dist(scale(rr40M))
rr40Fit <- hclust(rr40Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
ghb41TDM2 <- removeSparseTerms(ghb41TDM, sparse = 0.98)
ghb41M <- as.matrix(ghb41TDM2)
ghb41Matrix <- dist(scale(ghb41M))
ghb41Fit <- hclust(ghb41Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
wc42TDM2 <- removeSparseTerms(wc42TDM, sparse = 0.975)
wc42M <- as.matrix(wc42TDM2)
wc42Matrix <- dist(scale(wc42M))
wc42Fit <- hclust(wc42Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
gwb43TDM2 <- removeSparseTerms(gwb43TDM, sparse = 0.98)
gwb43M <- as.matrix(gwb43TDM2)
gwb43Matrix <- dist(scale(gwb43M))
gwb43Fit <- hclust(gwb43Matrix, method = "ward.D2")
```

```{r, echo=FALSE}
bo44TDM2 <- removeSparseTerms(bo44TDM, sparse = 0.98)
bo44M <- as.matrix(bo44TDM2)
bo44Matrix <- dist(scale(bo44M))
bo44Fit <- hclust(bo44Matrix, method = "ward.D2")
```

```{r, echo=FALSE, fig.height=6, fig.width=12}
plot(gw1Fit, main = "George Washington Term Dendrogram")
plot(ja2Fit, main = "John Adams Term Dendrogram")
plot(tj3Fit, main = "Thomas Jefferson Term Dendrogram")
plot(jm4Fit, main = "James Madison Term Dendrogram")
plot(jm5Fit, main = "James Monroe Term Dendrogram")
plot(jqa6Fit, main = "John Quincy Adams Term Dendrogram")
plot(aj7Fit, main = "Andrew Jackson Term Dendrogram")
plot(mvb8Fit, main = "Martin van Buren Term Dendrogram")
plot(jt10Fit, main = "John Tyler Term Dendrogram")
plot(jp11Fit, main = "James Polk Term Dendrogram")
plot(zt12Fit, main = "Zachary Taylor Term Dendrogram")
plot(mf13Fit, main = "Millard Fillmore Term Dendrogram")
plot(fp14Fit, main = "Franklin Pierce Term Dendrogram")
plot(jb15Fit, main = "James Buchanan Term Dendrogram")
plot(al16Fit, main = "Abraham Lincoln Term Dendrogram")
plot(aj17Fit, main = "Andrew Johnson Term Dendrogram")
plot(usg18Fit, main = "Ulysses S. Grant Term Dendrogram")
plot(rbh19Fit, main = "Rutherford B. Hayes Term Dendrogram")
plot(ca21Fit, main = "Chester A. Arthur Term Dendrogram")
plot(gc22Fit, main = "Grover Cleveland Term Dendrogram")
plot(bh23Fit, main = "Benjamin Harrison Term Dendrogram")
plot(wm25Fit, main = "William McKinley Term Dendrogram")
plot(tr26Fit, main = "Theodore Roosevelt Term Dendrogram")
plot(wt27Fit, main = "William H. Taft Term Dendrogram")
plot(ww28Fit, main = "Woodrow Wilson Term Dendrogram")
plot(wh29Fit, main = "Warren Harding Term Dendrogram")
plot(cc30Fit, main = "Calvin Coolidge Term Dendrogram")
plot(hh31Fit, main = "Herbert Hoover Term Dendrogram")
plot(fdr32Fit, main = "Franklin D. Roosevelt Term Dendrogram")
plot(ht33Fit, main = "Harry S. Truman Term Dendrogram")
plot(de34Fit, main = "Dwight D. Eisenhower Term Dendrogram")
plot(jfk35Fit, main = "John F. Kennedy Term Dendrogram")
plot(lbj36Fit, main = "Lyndon B. Johnson Term Dendrogram")
plot(rn37Fit, main = "Richard Nixon Term Dendrogram")
plot(gf38Fit, main = "Gerald R. Ford Term Dendrogram")
plot(jc39Fit, main = "Jimmy Carter Term Dendrogram")
plot(rr40Fit, main = "Ronald Reagan Term Dendrogram")
plot(ghb41Fit, main = "George H.W. Bush Term Dendrogram")
plot(wc42Fit, main = "William J. Clinton Term Dendrogram")
plot(gwb43Fit, main = "George W. Bush Term Dendrogram")
plot(bo44Fit, main = "Barack Obama Term Dendrogram")
```

## Word clouds

The following word clouds highlight the frequent terms used by every president and show how the State of the Union speeches evolved by presidency.

```{r, echo=FALSE, message=FALSE}
library(slam)
library(wordcloud)

gw1Freq <- row_sums(gw1TDM, na.rm=T)
gw1Word <- sort(gw1Freq, decreasing = T)

ja2Freq <- row_sums(ja2TDM, na.rm=T)
ja2Word <- sort(ja2Freq, decreasing = T)

tj3Freq <- row_sums(tj3TDM, na.rm=T)
tj3Word <- sort(tj3Freq, decreasing = T)

jm4Freq <- row_sums(jm4TDM, na.rm=T)
jm4Word <- sort(jm4Freq, decreasing = T)

jm5Freq <- row_sums(jm5TDM, na.rm=T)
jm5Word <- sort(jm5Freq, decreasing = T)

jqa6Freq <- row_sums(jqa6TDM, na.rm=T)
jqa6Word <- sort(jqa6Freq, decreasing = T)

aj7Freq <- row_sums(aj7TDM, na.rm=T)
aj7Word <- sort(aj7Freq, decreasing = T)

mvb8Freq <- row_sums(mvb8TDM, na.rm=T)
mvb8Word <- sort(mvb8Freq, decreasing = T)

jt10Freq <- row_sums(jt10TDM, na.rm=T)
jt10Word <- sort(jt10Freq, decreasing = T)

jp11Freq <- row_sums(jp11TDM, na.rm=T)
jp11Word <- sort(jp11Freq, decreasing = T)

zt12Freq <- row_sums(zt12TDM, na.rm=T)
zt12Word <- sort(zt12Freq, decreasing = T)

mf13Freq <- row_sums(mf13TDM, na.rm=T)
mf13Word <- sort(mf13Freq, decreasing = T)

fp14Freq <- row_sums(fp14TDM, na.rm=T)
fp14Word <- sort(fp14Freq, decreasing = T)

jb15Freq <- row_sums(jb15TDM, na.rm=T)
jb15Word <- sort(jb15Freq, decreasing = T)

al16Freq <- row_sums(al16TDM, na.rm=T)
al16Word <- sort(al16Freq, decreasing = T)

aj17Freq <- row_sums(aj17TDM, na.rm=T)
aj17Word <- sort(aj17Freq, decreasing = T)

usg18Freq <- row_sums(usg18TDM, na.rm=T)
usg18Word <- sort(usg18Freq, decreasing = T)

rbh19Freq <- row_sums(rbh19TDM, na.rm=T)
rbh19Word <- sort(rbh19Freq, decreasing = T)

ca21Freq <- row_sums(ca21TDM, na.rm=T)
ca21Word <- sort(ca21Freq, decreasing = T)

gc22Freq <- row_sums(gc22TDM, na.rm=T)
gc22Word <- sort(gc22Freq, decreasing = T)

bh23Freq <- row_sums(bh23TDM, na.rm=T)
bh23Word <- sort(bh23Freq, decreasing = T)

wm25Freq <- row_sums(wm25TDM, na.rm=T)
wm25Word <- sort(wm25Freq, decreasing = T)

tr26Freq <- row_sums(tr26TDM, na.rm=T)
tr26Word <- sort(tr26Freq, decreasing = T)

wt27Freq <- row_sums(wt27TDM, na.rm=T)
wt27Word <- sort(wt27Freq, decreasing = T)

ww28Freq <- row_sums(ww28TDM, na.rm=T)
ww28Word <- sort(ww28Freq, decreasing = T)

wh29Freq <- row_sums(wh29TDM, na.rm=T)
wh29Word <- sort(wh29Freq, decreasing = T)

cc30Freq <- row_sums(cc30TDM, na.rm=T)
cc30Word <- sort(cc30Freq, decreasing = T)

hh31Freq <- row_sums(hh31TDM, na.rm=T)
hh31Word <- sort(hh31Freq, decreasing = T)

fdr32Freq <- row_sums(fdr32TDM, na.rm=T)
fdr32Word <- sort(fdr32Freq, decreasing = T)

ht33Freq <- row_sums(ht33TDM, na.rm=T)
ht33Word <- sort(ht33Freq, decreasing = T)

de34Freq <- row_sums(de34TDM, na.rm=T)
de34Word <- sort(de34Freq, decreasing = T)

jfk35Freq <- row_sums(jfk35TDM, na.rm=T)
jfk35Word <- sort(jfk35Freq, decreasing = T)

lbj36Freq <- row_sums(lbj36TDM, na.rm=T)
lbj36Word <- sort(lbj36Freq, decreasing = T)

rn37Freq <- row_sums(rn37TDM, na.rm=T)
rn37Word <- sort(rn37Freq, decreasing = T)

gf38Freq <- row_sums(gf38TDM, na.rm=T)
gf38Word <- sort(gf38Freq, decreasing = T)

jc39Freq <- row_sums(jc39TDM, na.rm=T)
jc39Word <- sort(jc39Freq, decreasing = T)

rr40Freq <- row_sums(rr40TDM, na.rm=T)
rr40Word <- sort(rr40Freq, decreasing = T)

ghb41Freq <- row_sums(ghb41TDM, na.rm=T)
ghb41Word <- sort(ghb41Freq, decreasing = T)

wc42Freq <- row_sums(wc42TDM, na.rm=T)
wc42Word <- sort(wc42Freq, decreasing = T)

gwb43Freq <- row_sums(gwb43TDM, na.rm=T)
gwb43Word <- sort(gwb43Freq, decreasing = T)

bo44Freq <- row_sums(bo44TDM, na.rm=T)
bo44Word <- sort(bo44Freq, decreasing = T)

pal <- brewer.pal(8, "Dark2")
```

#### George Washington (1790-1796)
```{r, echo=FALSE}
wordcloud(words = names(gw1Word), freq = gw1Word, min.freq = 15, random.order = F, colors = pal)
```

#### John Adams (1797-1800)
```{r, echo=FALSE}
wordcloud(words = names(ja2Word), freq = ja2Word, min.freq = 10, random.order = F, colors = pal)
```

#### Thomas Jefferson (1801-1808)
```{r, echo=FALSE, warning=FALSE}
wordcloud(words = names(tj3Word), freq = tj3Word, min.freq = 30, random.order = F, colors = pal)
```

#### James Madison (1809-1816)
```{r, echo=FALSE}
wordcloud(words = names(jm4Word), freq = jm4Word, min.freq = 20, random.order = F, colors = pal)
```

#### James Monroe (1817-1824)
```{r, echo=FALSE}
wordcloud(words = names(jm5Word), freq = jm5Word, min.freq = 30, random.order = F, colors = pal)
```

#### John Quincy Adams (1825-1828)
```{r, echo=FALSE}
wordcloud(words = names(jqa6Word), freq = jqa6Word, min.freq = 20, random.order = F, colors = pal)
```

#### Andrew Jackson (1829-1836)
```{r, echo=FALSE}
wordcloud(words = names(aj7Word), freq = aj7Word, min.freq = 42, random.order = F, colors = pal)
```

#### Martin van Buren (1837-1840)
```{r, echo=FALSE}
wordcloud(words = names(mvb8Word), freq = mvb8Word, min.freq = 30, random.order = F, colors = pal)
```

#### John Tyler (1841-1844)
```{r, echo=FALSE}
wordcloud(words = names(jt10Word), freq = jt10Word, min.freq = 20, random.order = F, colors = pal)
```

#### James Polk (1845-1848)
```{r, echo=FALSE}
wordcloud(words = names(jp11Word), freq = jp11Word, min.freq = 37, random.order = F, colors = pal)
```

#### Zachary Taylor (1849)
```{r, echo=FALSE}
wordcloud(words = names(zt12Word), freq = zt12Word, min.freq = 8, random.order = F, colors = pal)
```

#### Millard Fillmore (1850-1852)
```{r, echo=FALSE}
wordcloud(words = names(mf13Word), freq = mf13Word, min.freq = 20, random.order = F, colors = pal)
```

#### Franklin Pierce (1853-1856)
```{r, echo=FALSE}
wordcloud(words = names(fp14Word), freq = fp14Word, min.freq = 20, random.order = F, colors = pal)
```

#### James Buchanan (1857-1860)
```{r, echo=FALSE}
wordcloud(words = names(jb15Word), freq = jb15Word, min.freq = 30, random.order = F, colors = pal)
```

#### Abraham Lincoln (1861-1864)
```{r, echo=FALSE}
wordcloud(words = names(al16Word), freq = al16Word, min.freq = 15, random.order = F, colors = pal)
```

#### Andrew Johnson (1865-1868)
```{r, echo=FALSE}
wordcloud(words = names(aj17Word), freq = aj17Word, min.freq = 20, random.order = F, colors = pal)
```

#### Ulysses S. Grant (1869-1876)
```{r, echo=FALSE}
wordcloud(words = names(usg18Word), freq = usg18Word, min.freq = 30, random.order = F, colors = pal)
```

#### Rutherford B. Hayes (1877-1880)
```{r, echo=FALSE}
wordcloud(words = names(rbh19Word), freq = rbh19Word, min.freq = 30, random.order = F, colors = pal)
```

#### Chester A. Arthur (1881-1884)
```{r, echo=FALSE}
wordcloud(words = names(ca21Word), freq = ca21Word, min.freq = 15, random.order = F, colors = pal)
```

#### Grover Cleveland (1885-1888, 1893-1896)
```{r, echo=FALSE}
wordcloud(words = names(gc22Word), freq = gc22Word, min.freq = 45, random.order = F, colors = pal)
```

#### Benjamin Harrison (1889-1892)
```{r, echo=FALSE}
wordcloud(words = names(bh23Word), freq = bh23Word, min.freq = 32, random.order = F, colors = pal)
```

#### William McKinley (1897-1900)
```{r, echo=FALSE}
wordcloud(words = names(wm25Word), freq = wm25Word, min.freq = 30, random.order = F, colors = pal)
```

#### Theodore Roosevelt (1901-1908)
```{r, echo=FALSE}
wordcloud(words = names(tr26Word), freq = tr26Word, min.freq = 125, random.order = F, colors = pal)
```

#### William H. Taft (1909-1912)
```{r, echo=FALSE}
wordcloud(words = names(wt27Word), freq = wt27Word, min.freq = 50, random.order = F, colors = pal)
```

#### Woodrow Wilson (1913-1920)
```{r, echo=FALSE}
wordcloud(words = names(ww28Word), freq = ww28Word, min.freq = 37, random.order = F, colors = pal)
```

#### Warren Harding (1921-1922)
```{r, echo=FALSE}
wordcloud(words = names(wh29Word), freq = wh29Word, min.freq = 17, random.order = F, colors = pal)
```

#### Calvin Coolidge (1923-1928)
```{r, echo=FALSE}
wordcloud(words = names(cc30Word), freq = cc30Word, min.freq = 33, random.order = F, colors = pal)
```

#### Herbert Hoover (1929-1932)
```{r, echo=FALSE}
wordcloud(words = names(hh31Word), freq = hh31Word, min.freq = 30, random.order = F, colors = pal)
```

#### Franklin D. Roosevelt (1934-1945)
```{r, echo=FALSE}
wordcloud(words = names(fdr32Word), freq = fdr32Word, min.freq = 25, random.order = F, colors = pal)
```

#### Harry S. Truman (1946-1953)
```{r, echo=FALSE, warning=FALSE}
wordcloud(words = names(ht33Word), freq = ht33Word, min.freq = 65, random.order = F, colors = pal)
```

#### Dwight D. Eisenhower (1953-1961)
```{r, echo=FALSE, warning=FALSE}
wordcloud(words = names(de34Word), freq = de34Word, min.freq = 90, random.order = F, colors = pal)
```

#### John F. Kennedy (1961-1963)
```{r, echo=FALSE}
wordcloud(words = names(jfk35Word), freq = jfk35Word, min.freq = 15, random.order = F, colors = pal)
```

#### Lyndon B. Johnson (1964-1969)
```{r, echo=FALSE}
wordcloud(words = names(lbj36Word), freq = lbj36Word, min.freq = 35, random.order = F, colors = pal)
```

#### Richard Nixon (1970-1974)
```{r, echo=FALSE}
wordcloud(words = names(rn37Word), freq = rn37Word, min.freq = 20, random.order = F, colors = pal)
```

#### Gerald R. Ford (1975-1977)
```{r, echo=FALSE}
wordcloud(words = names(gf38Word), freq = gf38Word, min.freq = 23, random.order = F, colors = pal)
```

#### Jimmy Carter (1978-1981)
```{r, echo=FALSE, warning=FALSE}
wordcloud(words = names(jc39Word), freq = jc39Word, min.freq = 90, random.order = F, colors = pal)
```

#### Ronald Reagan (1982-1988)
```{r, echo=FALSE}
wordcloud(words = names(rr40Word), freq = rr40Word, min.freq = 45, random.order = F, colors = pal)
```

#### George H.W. Bush (1989-1992)
```{r, echo=FALSE}
wordcloud(words = names(ghb41Word), freq = ghb41Word, min.freq = 15, random.order = F, colors = pal)
```

#### William J. Clinton (1993-2000)
```{r, echo=FALSE}
wordcloud(words = names(wc42Word), freq = wc42Word, min.freq = 42, random.order = F, colors = pal)
```

#### George W. Bush (2001-2008)
```{r, echo=FALSE}
wordcloud(words = names(gwb43Word), freq = gwb43Word, min.freq = 30, random.order = F, colors = pal)
```

#### Barack Obama (2009-2015)
```{r, echo=FALSE}
wordcloud(words = names(bo44Word), freq = bo44Word, min.freq = 38, random.order = F, colors = pal)
```
